Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-04-23 doc/GMP2015/revision_2/s01_introduction.tex
Working on the paper revision.

% \section{Response to reviewers}

% \cutout{This means} the text was removed.
% \addin{This means} the text was added.
% \notetoreviewer{This means} it is a note to the reviewer, and would not appear in real document.
% % \movedfrom{This means} the text was moved from another part of the document.

% \subsection{Reviewer 1}
% No suggestions made.

% \subsection{Reviewer 2}
% % \noindent
% % Changes made:
% \begin{itemize}
%   \item We added a new example in figures to illustrate definitions/proofs/comparisons. And we added a new conjecture relating this method to the Modified Frontier Algorithm, the most recent and best current algorithm.
%   \item Abstract was shortened and clarified.
%   \item Basic definitions from Section \ref{sec:prelim} were moved to the new \ref{sec:appendix:defs}. Definition of qusecs was made more prominent, left in Section \ref{sec:prelim}. Slight modification to definition of DR-plan.
%   \item Subsections \ref{sec:cont} and \ref{sec:intro:prevwork} were flipped. This gives readers an idea of the contributions before they read about the current state of the field.
%   \item Section \ref{sec:DRP} had some proofs moved back in from the appendix. Section \ref{sec:recomb} had a proof moved back in from the appendix. Section \ref{sec:conclusion} had future work moved back in from the appendix.
%   \item Numerous grammatical, syntactical, and spelling errors were fixed. Made hyphen usage consistent.
%   \item Images were properly aligned and cropped.
% \end{itemize}

% \subsection{Reviewer 3}
% We emailed the co-chairs, and they replied that it would fit in the conference, leaving it at our discretion to withdraw. We have decided to continue.

@@

Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often self-similar~\cite{2012arXiv1204.6389G} spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and consists of a few repeated motifs appearing in disordered arrangements. Each layer is not necessarily planar, i.e., it consists of multiple, inter-constraining planar (genus 0) monolayers. Furthermore, a layer is often  either
\vemph{isostatic or underconstrained} (\vemph{not self-stressed},\seedefsd). A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.

\noindent
\textbf{Note on Scope:} In this manuscript we only study finite 2D structures. \dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.

@@

Examples of such materials (See Figure \ref{fig:material_examples})
include:
%
\begin{enumerate}
    \item \label{materialexample1} Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.

@@

    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski}, bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.

    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and materials that behave like assemblies of 2D particles under non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}. See Section \ref{sec:bodypin}.
\end{enumerate}
%
In order to study structural and mechanical properties of a material layer, it is natural to model a material layer as a solution or realization of a geometric constraint system of appropriate types of geometric primitives, under metric or algebraic constraints. Such 2D \dfn{qusecs}, quasi-uniform or self-similar constraint systems \seedefsprelim, can be used to understand or design material layers (their solutions) with desired properties.

\subsection{Contributions and Organization}
\label{sec:cont}

The contributions of this paper are the following.
\begin{itemize}
  \item In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan. This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.

  \item In Section \ref{sec:DRP}, we navigate the NP-hardness barrier (discussed in the following subsection), for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.

  \item Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}). The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).

  \item In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}. Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components. When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).

  \item In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}. For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems. For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).

  \item We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.
\end{itemize}

\subsection{Previous Work on Relevant 2D Geometric Constraint Systems}
\label{sec:intro:prevwork}

We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.

\medskip\noindent
\header{(i) Finding (vertex)-maximal generically rigid subsystems}
Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs. When the input itself is rigid, these algorithms do nothing, i.e., compute the identity function.

However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as \seedefs\ \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.

\medskip\noindent
\header{(ii) Optimal Recursive Decomposition (DR-planning).}
Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system. Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.

However, for general 2D qusecs that could be overconstrained, even when restricted to bar-joint systems, the optimal DR-planning problem was shown to be NP-hard \cite{lomonosov2004graph}.

\medskip\noindent
\header{(iii) DR-plans for special classes and with other
criteria.}
For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal. This satisfies the so-called \dfn{Church-Rosser} criterion, leading to highly efficient DR-planning algorithms. For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system. \seedefsc.

While polynomial time algorithms were given to generate DR-plans meeting the cluster minimality criterion \cite{hoffman2001decompositionI}, no such algorithm is known for the latter criterion.

\medskip\noindent
\header{(iv) Optimal Recombination and Solution Space Navigation.}
For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan. This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}. Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan. For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.

However, the generality of these algorithms trades-off against efficiency, and despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan. They are prohibitively slow in practice. We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.

\medskip\noindent
\header{(v) Configuration Spaces of Underconstrained Systems.}
For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths. These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs. While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.

% \medskip\noindent
% \note All proofs appear in the Appendix.
% \medskip

>>>>

@@

\todo{Main overview/intro.}
Geometric constraint systems are ubiquitous in material and mechanical engineering, robotics, bioinformatics, and many other fields. However, the design and analysis of these systems is made difficult by i

\subsection{Recursive Decomposition Problem}
The first problem we consider is the optimal recursive decomposition problem of highly general geometric constraint systems with the sole requirement that local rigidity is characterized as a sparsity condition of an underlying (hyper)graph. We call these systems \nameit. This characterization has broad applicability in practice, and we provide several concrete examples.

Such decomposition is crucial for realizing and designing the geometric constraint system and its properties such as rigidity, stability, and stress response to external loads. In the literature, this problem is called the optimal decomposition recombination (DR-) planning problem, which, in general, is NP-hard. Therefore, previous algorithms for the problem did not consider optimality and instead focused on other desirable criteria of the output DR-plan. By concentrating on geometric constraint system that are not overconstrained, we give a polynomial time algorithm for optimal DR-planning by capturing the notion of a canonical DR-plan; we show that all canonical DR-plans are optimal for systems that are not overconstrained. This immediately gives a polynomial time algorithm for obtaining a DR-planning that is not only optimal but also has the previously studied desirable properties. In fact, we conjecture that one of the earlier algorithms also produces an optimal DR-plan and that every optimal DR-plan is in some formal sense canonical.

% Knowing an optimal DR-plan is crucial when working with isostatic geometric constraint systems with an underlying sparsity condition \nameit. It allows for efficient design and analysis of these systems with various desirable properties. The importance of an optimal plan is particularly evident when these systems \nameit are self-similar or quasi-uniform, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}, as the complexity of the plan is even further reduced.

When realizing such a system \nameit, the task would require finding real solutions to a large multivariate polynomial system (of inequalities and equalities representing the constraints). Without DR-planning, this requires double exponential time in the number of variables (even if orientation type is specified). With DR-planning, the complexity is dominated by the size of the largest subsystem that is solved, or recombined from the solutions of its child subsystems, i.e., the maximum fan-in occurring in a DR-plan. Thus, finding an optimal DR-plan is important for realization.

\subsection{Recombination Problem}
\todo{The second problem we consider...}
We additionally use the notion of Cayley complexity to quantify the difficulty of each step of recombination. Once a subsystem is found that has a convex Cayley configuration space, the space can be efficiently searched to find a realization that satisfies the additional constraints of the original system.
This allows for even greater reduction of the complexity, by realizing large, indecomposable systems in a way that avoids working with large systems of equations.

\subsection{Introducing Qusecs}
When working on these problems, there is a large, natural class of constraint systems that we call \dfn{qusecs}, a contraction of ``quasi-uniform or self-similar constraint systems''.
%
Some properties of natural and engineered materials can be analyzed by treating them as two dimensional (2D) layers. As illustrated by the examples below, the structure within each layer is often: self-similar\footnote{In this manuscript we only study finite 2D structures. \dfn{Self-similarity} refers to the result of finitely many levels of hierarchy or subdivision in an iterated scheme to generate self-similar structures.}~\cite{2012arXiv1204.6389G}, spanning multiple scales; generally aperiodic and quasi-uniform within any one scale; and composed of a few repeated motifs appearing in disordered arrangements.
Note that a layer is not necessarily planar, it can consist of multiple, inter-constraining planar (genus 0) monolayers. Furthermore, a layer is often  either \vemph{isostatic} or \vemph{underconstrained} (\vemph{not self-stressed}, \seedefsd). A variety of quasi-uniform (aperiodic) and self-similar, layered material structures are a natural consequence of design objectives such as rigidity or flexibility, minimizing mass, optimally distributing external stresses and participating in the assembly of diverse and multifunctional, larger  structures.
%
The importance of an optimal DR-plan is particularly evident when a system is a qusecs. The quasi-uniform or self-similar properties mean that solutions for one subsystem can solve other subsystems, thus causing further reduction in the complexity of the plan. This is shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.

@@

% Examples of such materials (See Figure \ref{fig:material_examples}) include:
% In order to study structural and mechanical properties of a material layer, it is natural to model a material layer as a solution or realization of a geometric constraint system of appropriate types of geometric primitives, under metric or algebraic constraints.
% Such 2D \dfn{qusecs}, quasi-uniform or self-similar constraint systems \seedefsprelim, can be used to understand or design material layers (their solutions) with desired properties.
Some materials that are readily modeled as qusecs include:
%
\begin{enumerate}
    \item \label{materialexample1} Cross-sections of microtubule structures~\cite{microtubule_necklace} (Figure \ref{fig:material_examples:microtubule}) e.g., in ciliary membranes and transitions~\cite{microtubule_cilia}.

@@

    \item \label{materialexample4} More recent, engineered examples, including disordered graphene layers~\cite{Graphene1}~\cite{Graphene2} sometimes reinforced by microfibrils; and DNA assemblies including a recent Sierpinski gasket~\cite{self_assembly_sierpinski} (Figure \ref{fig:material_examples:sierpinski}), bringing other self-similar structures~\cite{wikimediacommons2012subdivision} within reach.

    \item \label{materialexample5} Silica bi-layers~\cite{silica_bilayers}, glass~\cite{sructure_of_2d_glass}, and materials that behave like assemblies of 2D particles under non-overlap constraints, i.e, like jammed disks on the plane~\cite{jammed_disks}. See Section \ref{sec:bodypin}.
\end{enumerate}

% \subsection{old}

\subsection{Organization and Contributions}
\label{sec:cont}

\todo{In Section \ref{sec:prelim}, we give definitions and give the new notion of qusecs and the canonical/optimal DR-plan. This is relevant to Examples \ref{materialexample1} and \ref{materialexample2}, where we model them as bar-joint systems and discuss achieving isostaticity, distribution of stresses in self-similar, and other important concepts.}

% In Section \ref{sec:DRP}, we navigate the NP-hardness barrier (discussed in the following subsection), for finding optimal DR-plans by defining a so-called \dfn{canonical} DR-plan and showing a strong Church-Rosser property: \vemph{all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal}.

% Also in Section \ref{sec:DRP}, we give an efficient (\candrpcomplexity) algorithm to find a canonical (and hence optimal) DR-plan for all 3 types of 2D qusecs mentioned above (Sections \ref{sec:DRP}, \ref{sec:bodypin}, and \ref{sec:pinnedline}).
% The canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems.
% Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}).

In Section \ref{sec:DRP}, we define a so-called \dfn{canonical DR-plan} and prove a strong Church-Rosser property: all canonical DR-plans for isostatic or underconstrained 2D qusecs are optimal. In so doing, we navigate the NP-hardness barrier present in the general form of the DR-planning problem; the canonical DR-plan elucidates the essence of the NP-hardness of finding optimal DR-plans for over-constrained systems. Furthermore, our optimal/canonical DR-plan satisfies desirable properties such as the previously studied Cluster Minimality \cite{hoffman2001decompositionI} (see Figure~\ref{fig:demo_graph:clustmindrp}). Also in this section, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.

% that such a plan will be optimal, in the sense that it minimizes the fan-in over all nodes, for isostatic bar-joint graphs.
% Additionally, a polynomial time (\candrpcomplexity) algorithm is provided to find a canonical DR-plan for isostatic bar-joint graphs.

In Section \ref{sec:recomb}, we give a method to deal with the algebraic complexity of recombining the realizations or solutions of child subsystems into a solution of the parent system \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}. Specifically, we define the problem of minimally modifying the indecomposable recombination system so that it becomes decomposable via a small DR-plan and yet preserves the original solutions in an efficiently searchable manner. In Section \ref{sec:table}, we show formal connection to well known problems such as optimal completion of underconstrained systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree} and to find paths within the connected components. When the modifications are bounded, we obtain new, efficient algorithms for realizing both isostatic and underconstrained qusecs by leveraging recent results about Cayley parameters in \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} (see Sections \ref{sec:2-tree-reduction} and \ref{sec:tdecomp}).

\todo{Section \ref{sec:recomb} addresses the issue of large, indecomposable subgraphs in the optimal DR-plan of bar-joint graphs by proposing a novel method of modification. By dropping edges to get a convex configuration space, realizations of the original linkage can be found. Also problem relations...}

While both of these sections focus on bar-joint graphs, the theory is easily extended to other qusecs. The following sections focus on explicit applications by modeling rigidity in certain materials as qusecs.

In Section \ref{sec:bodypin} and \ref{sec:pinnedline}, we briefly describe applications of the above techniques to modeling, analyzing, and designing specific properties in 2D material layers~\cite{Jackson2008bodypin}. For Examples \ref{materialexample4} and \ref{materialexample5}, we discuss boundary-conditions for achieving various desired properties of body-hyperpin systems. For Example \ref{materialexample3}, we discuss canonical and optimal DR-plans for pinned line incidence systems \cite{sitharam2014incidence}).

We intend to make software implementation and videos available upon request and publicly available for the final version of the paper.

\subsection{The Importance of an Optimal DR-plan}
% \subsection{Importance of an Optimal DR-plan}
% The use of a DR-plan in designing especially self-similar qusecs is immediately evident, shown in Figures \ref{fig:c2c3ofk33s} and \ref{fig:bodypindrp}.

% In fact, knowing an optimal DR-plan is crucial for design and analysis of even non-self-similar qusecs and their realizations with various desirable properties.

Furthermore, recursive block decomposition of the rigidity matrix, stress space, and external stresses can be obtained using the DR-plan. Each block is the rigidity matrix of an isostatic subgraph. The blocks intersect on isostatic or trivial subgraphs. This recursive block decomposition of the rigidity matrix $R$ corresponds to a recursive decomposition of the stress vectors $s$ that balance a given external stress $t$ since $sR = t$. Moreover, the external stress vector $t$ is recursively ``distributed'' into external stresses for each node of the DR-plan. When the rigidity matrix and stress vectors contain indeterminates, this can be used to design underlying graph and parameters $\delta$ to obtain an optimal distribution of stresses on the geometric primitives. When the realization of the system \nameit is given, this can be used to analyze the distribution of a given external load.

### 2015-04-23 doc/GMP2015/revision_2/s03_drp.tex
Working on the paper revision.

\sidenote{In this section, $G$ is always assumed to be isostatic.}

@@

% \subsection{Algorithm}

>>>>

\subsection{Previous Work}
\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}

\begin{enumerate}
    \item
    \header{Finding (vertex)-maximal generically rigid subsystems}
    Fast, graph-based algorithms exist (pebble-game \cite{Jacobs:1997:PG,lee2005finding}), for locating all maximal, \dfn{generically rigid} subsystems \seedefs. When the input itself is rigid, these algorithms do nothing, i.e., compute the identity function.

    However, both for self-similar or just aperiodic 2D qusecs, it is imperative to recursively decompose rigid systems into their rigid subsystems, down to the level of geometric primitives, in order to understand or design properties at all scales, such as \seedefs\ \dfn{rigidity}, \dfn{flexes}, distribution of \dfn{external stresses}, boundary conditions for \dfn{isostaticity}, as well as behavior under constraint variations.

    \item
    \header{Optimal Recursive Decomposition (DR-planning).}
    Recursive decomposition of geometric constraint systems has been formalized \cite{hoffman2001decompositionI,hoffman2001decompositionII} and well-studied \cite{jermann2006decomposition,sitharam2005combinatorial} as the \dfn {Decomposition-Recombination (DR-) planning} problem \seedefsprelim. For the abovementioned classes of 2D qusecs, generic rigidity is a combinatorial property and hence each level of the decomposition should, in principle, be achievable by a graph-based algorithm as in (1), without involving geometric information in the constraint system. Since many such decompositions can exist for a given constraint system, criteria defining desirable or optimal DR-plans and DR-planning algorithms were given in \cite{hoffman2001decompositionI}. An \dfn{optimal DR-plan} is one that minimizes the \dfn{size} \seedefsprelim, i.e., the maximum number of child subsystems of any parent system. Being exponential in the size, the complexity of solving the parent constraint system is overwhelmingly dominated by the complexity of solving the child systems.

    However, for general 2D qusecs that could be overconstrained, even when restricted to bar-joint systems, the optimal DR-planning problem was shown to be NP-hard \cite{lomonosov2004graph}.

    \item
    \header{DR-plans for special classes and with other criteria.}
    For a special class of 2D qusecs, namely \dfn{tree-decomposable} systems  \cite{fudos1997graph,owen1991algebraic,joan-arinyo2004revisiting}  common in computer aided mechanical design, (which includes ruler-and-compass and Henneberg-I constructible systems), all DR-plans turn out to be optimal. This satisfies the so-called \dfn{Church-Rosser} criterion, leading to highly efficient DR-planning algorithms. For general 2D qusecs, alternate criteria were suggested such as \dfn{cluster minimality} requiring parent systems to be composed of a minimal set of at least 2 rigid proper subsystems (i.e., no proper subset forms a rigid system); and \dfn{proper maximality}, requiring child subsystems to be maximal rigid proper subsystems of the parent system. \seedefsc.

    \indent
    While polynomial time algorithms were given to generate DR-plans meeting the cluster minimality criterion \cite{hoffman2001decompositionI}, no such algorithm is known for the latter criterion.
\end{enumerate}

\subsection{Theory}
\sidenote{In this section, $G$ is always assumed to be isostatic.}

@@

\subsection{Algorithm}

### 2015-04-23 doc/GMP2015/revision_2/s04_recombination.tex
Working on the paper revision.

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}
%
In this section, we consider the \dfn{optimal recombination} problem of combining specific solutions of subsystems in a DR-plan into a solution of their parent system $I$ (without loss of generality, at the top level of the DR-plan). In the case of isostatic qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we seek \vemph{solution(s) (among a finite large number of solutions) with a specific orientation or chirality}. In the case of underconstrained qusecs the subsystems are the multiple roots of the DR-plan, the parent system $I$ is underconstrained, and we typically seek an efficient algorithmic description of \vemph{connected component(s) of solutions with a specific orientation or chirality}.

@@

\medskip\noindent
\header{(a) Small DR-plan.}
We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can potentially be solved or realized easily once the orientation type of the solution is known  (when the DR-plan size is small enough).

\medskip\noindent
\header{(b) Solution preservation.}
Moreover, the union of solution spaces of the systems in the family $I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.

\medskip\noindent
\header{(c) Efficient search.}
Finally, this union can be projected into the so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type. For the case when the modification is bounded, this approach provides an efficient algorithm for recombination. We first define the decision version of the problem of optimal modification for decomposition. The standard optimization versions are straightforward.

@@

>>>>

% \section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\section{Recombination and Problem Relationships}
\label{sec:recomb}
%
In this section, we consider the \dfn{optimal recombination} problem of combining specific solutions of subsystems in a DR-plan into a solution of their parent system $I$ (without loss of generality, at the top level of the DR-plan). In the case of isostatic qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we seek \vemph{solution(s) (among a finite large number of solutions) with a specific orientation or chirality}. In the case of underconstrained qusecs the subsystems are the multiple roots of the DR-plan, the parent system $I$ is underconstrained, and we typically seek an efficient algorithmic description of \vemph{connected component(s) of solutions with a specific orientation or chirality}.

@@

\begin{enumerate}[(a)]
    \item
    \header{Small DR-plan.}
    We obtain a  parameterized family of systems $I_{\lambda_F}$---one for each value $\lambda_F$ for the parameters $F$,  all of which have small DR-plans. Thus, given a value $v$ for $\lambda_F$, the system $I_v$ can potentially be solved or realized easily once the orientation type of the solution is known  (when the DR-plan size is small enough).

    \item
    \header{Solution preservation.}
    Moreover, the union of solution spaces of the systems in the family $I_{\lambda_F}$ is guaranteed to contain all of $I$'s solutions.

    \item
    \header{Efficient search.}
    Finally, this union can be projected into the so-called \dfn{Cayley} or distance parameter space $\lambda_F$ that is convex or otherwise easy to traverse in order to search for $I$'s solution (or connected component) of the desired orientation type. For the case when the modification is bounded, this approach provides an efficient algorithm for recombination. We first define the decision version of the problem of optimal modification for decomposition. The standard optimization versions are straightforward.
\end{enumerate}

@@

\subsection{Previous Work}
\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}

\begin{enumerate}
    \item
    \header{Optimal Recombination and Solution Space Navigation.}
    For the entire DR-plan, finding all desired solutions is barely tractable even if recombination of solved subsystems is easy for each indecomposable parent system in the DR-plan. This is because even for the simplest, highly decomposable systems with small DR-plans, the problem of finding even a single solution to the input system at the root of the DR-plan is NP-hard  \cite{saxe1979embeddability} and there is a combinatorial explosion of solutions \cite{borcea2004number}. Typically, however, the desired solution has a given orientation type, in which case, the crux of the difficulty is concentrated in the algebraic complexity of (re)combining child system solutions to give a solution to the parent system at any given level of the DR-plan. For fairly general 3D constraint systems, there are algorithms with formal guarantees that uncover underlying matroids to combinatorially obtain an optimal parameterization to minimize the algebraic complexity of the indecomposable parent (sub)systems that occur in the DR-plan \cite{sitharam2010optimized,sitharam2006well,sitharam2010reconciling}, provided the DR-plan meets some of the abovementioned criteria.

    However, the generality of these algorithms trades-off against efficiency, and despite the optimization, the best algorithms can still take exponential time in the number of child subsystems (which can be arbitrarily large even for optimal DR-plans) in order to guarantee all solutions of a given orientation type, even for a single (sub)system in a DR-plan. They are prohibitively slow in practice. We note that, utilizing the DR-plan and optimal recombination as a principled basis, high performance heuristics and software exists \cite{sitharam2006solution} to tame combinatorial explosion via user intervention.

    \item
    \header{Configuration Spaces of Underconstrained Systems.}
    For underconstrained 2D bar-joint and body-hyperpin qusecs obtained from various subclasses of tree-decomposable systems, algorithms have been developed to complete them into isostatic systems \cite{joan-arinyo2003transforming,sitharam2005combinatorial,gao2006ctree,sitharam2010convex} and to find paths within the connected components \cite{sitharam2011cayleyI,hidalgo2011reachability} of standard Cartesian configuration spaces. Most of the algorithms with formal guarantees leverage Cayley configuration space theory \cite{sitharam2010convex,sitharam2011cayleyI,sitharam2011cayleyII} to characterize subclasses of graphs and additional constraints that control combinatorial explosion, and provide faithful bijective representation of connected components and paths. These algorithms have decreasing efficiency as the subclass of systems gets bigger, with highest efficiency for underlying partial 2-tree graphs (alternately called tree-width 2, series-parallel, and $K_4$ minor avoiding), moderate efficiency for 1 degree-of-freedom (dof) graphs with low Cayley complexity (which include common linkages such as the Strandbeest, Limacon and Cardioid), and decreased efficiency for general 1-dof tree-decomposable graphs. While software suites exist  \cite{keycurriculum1995geometer,porta2014open,siemens1999d,todd2007geometry}, no such formal algorithms and guarantees are known for non-tree-decomposable systems.
\end{enumerate}
